---
title: "CKME136 - CAPSTONE - Exploration & Analysis"
author: "Mohammed Amir"
date: "April 17, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Step 1. Data Load Process
Dinesafe is an open dataset from City of Toronto Food Hygiene Inspection Report for the year 2015 and 2016.
Address is a full address dataset extracted from google map for the Dinesafe food premises using google geocode.
Load Dinesafe and Address Datasets, convert NULL values to NA. 
```{r , results='hide'}
Dinesafe = read.csv("D:/CAPSTONE/data/DineSafe_02162017.csv", na.strings='NULL')
Address  = read.csv("D:/CAPSTONE/data/ADDRESS_02262017.csv", na.strings='NULL')
```

Step 2. Dataset Exploration Process
2.1 - Identify the column names for each datasets
```{r , results='hide'}
## List Column names
colnames(Dinesafe)
cat("\n")
colnames(Address)
```
2.2 - Identify database dimensions, Address has 7 columns and Dinesafe has 17 columns
```{r , results='hide'}
## Review dimension of dataset (Row by column)
dim(Dinesafe)
cat("\n")
dim(Address)
```
2.3 - Summarise the datasets. Find the min, max, median for quartile quantitative values, as well as identify word category counts for the categorical values.
```{r , results='hide'}
## Review dataset summary
summary(Dinesafe)
cat("\n")
summary(Address)
```

2.4 - Identify the dataset structure such as int, factor, num for both datasets
```{r , results='hide'}
## Review dataset structure
str(Dinesafe)
cat("\n")
str(Address)
```

2.5 - Display top 5 sample data 
```{r , results='hide'}
head(Dinesafe,5)
head(Address,5)
```

Step 3 - Merge Dinesafe and Address datasets based on establishment id column
```{r , results='hide'}
Dinesafe <- merge(Dinesafe,Address,by="ESTABLISHMENT_ID")
```

3.1 - Analyise merged dataset
```{r , results='hide'}
dim(Dinesafe)     ## Identify dimension
cat("\n")
str(Dinesafe)     ## Identify structure
cat("\n")             
table(Dinesafe$CUISINE_TYPE, useNA = "always")                 ## Identify Cuisine Type
cat("\n")  
table(Dinesafe$ESTABLISHMENT_STATUS, useNA = "always")         ## Identify Review Rating
cat("\n")  
table(Dinesafe$DISTRICT, useNA = "always")                     ## Identify Districts
cat("\n")  
table(Dinesafe$SEVERITY)                                       ## Identify Severity Type

```
Step 4 -  Data Munging Step
4.1 - Data Cleaning and Transforming raw data into usable dataset.
```{r , results='hide'}
### Remove COURT_OUTCOME, AMOUNT_FINED & INFRACTION_DETAILS Columns from Dinesafe dataset
Dinesafe <- subset(Dinesafe, select = -c(ROW_ID, COURT_OUTCOME,AMOUNT_FINED,LONG_ADDRESS, INFRACTION_DETAILS) )
cat("\n")
### Remove doublicate Establishement Name and Address from dataset 
Dinesafe <- subset(Dinesafe, select = -c(ESTABLISHMENT_NAME.y, ESTABLISHMENT_ADDRESS) )
cat("\n")
## Rename ESTABLISHMENT_NAME.x column name to ESTABLISHMENT_NAME
colnames(Dinesafe)[colnames(Dinesafe) == 'ESTABLISHMENT_NAME.x'] <- 'ESTABLISHMENT_NAME'
```
# Plot missingness map using Amelia package
```{r , results='hide'}
#Quantify missing values
apply(Dinesafe, 2, function(x) sum(is.na(x)))
cat("\n")

# Plot missingness map using Amelia package
library(Amelia)    
missmap(Dinesafe, col = c("Red","Yellow"), y.cex = 0.8, x.cex = 0.8, legend = TRUE, rank.order = "False" ,main = "Dinesafe missingness map", y.labels = NULL,y.at = NULL)

```

4.3 - Change ACTION from factor to character to avoid error during imputation.
```{r , results='hide'}
## Convert Action column from factor to character type
Dinesafe$ACTION = as.character(Dinesafe$ACTION)
```

4.4 - Set catagorical level for Establishment status and Sevrity columns
```{r , results='hide'}
## Set Categorical Data Type Level for Establishment Status column
Dinesafe$ESTABLISHMENT_STATUS = factor(Dinesafe$ESTABLISHMENT_STATUS,levels=c("Closed","Conditional Pass", "Pass"))
cat("\n")

## Set Categorical Data Type Level for Severity column
Dinesafe$SEVERITY <- factor(Dinesafe$SEVERITY, levels = c("NA - Not Applicable", "N - No Action", "M - Minor", "S - Significant", "C - Crucial"))
```

4.5 - Describe quantitative values in Reveiw and Rate columns
```{r , results='hide'}
library(Hmisc)
## Describe Review data
describe(Dinesafe$REVIEW)
cat("\n")

## Describe Value data
describe(Dinesafe$VALUE)
cat("\n")
```



4.6 - Show complete rows from dataset
```{r , results='hide'}
## Complete Case Rows with no missing (NA) value
Complete_Dinesafe  <- Dinesafe[complete.cases(Dinesafe),]
nrow(Complete_Dinesafe)
```
4.7 - Impute NA values in REVIEW column based on mean value of each cuisine type
```{r , results='hide'}
## Impute Dinesafe$REVIEW with Mean Review Value for each missing review value based cuisine type
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)

```

4.7 - Impute NA values in VALUE column based on mean value of each cuisine type
```{r , results='hide'}
## Impute Dinesafe$VALUE with Mean Value for each missing value based cuisine type
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)
```

4.8 - Impute missing severity and action columns where establishment status is PASS
```{r , results='hide'}
## Impute Severity column if it is NA and Establishment Status is PASS
Dinesafe$SEVERITY[is.na(Dinesafe$SEVERITY) & Dinesafe$ESTABLISHMENT_STATUS == "Pass"] = "NA - Not Applicable"
cat("\n")
## Impute Action column if it is NA and Establishment Status is PASS & Severity is No Action
Dinesafe$ACTION[is.na(Dinesafe$ACTION) & Dinesafe$ESTABLISHMENT_STATUS == "Pass" & Dinesafe$SEVERITY == "NA - Not Applicable"] = "No Action Required"

```
4.9 - Check for incomplete rows
```{r , results='hide'}
## Check for non complete case
Dinesafe_NA  <- Dinesafe[!complete.cases(Dinesafe),]
cat("\n")
nrow(Dinesafe_NA)
```


Step 5 - Data Exploratory Analysis and Visualization
5.1 Univarient Data Analysis
```{r , results='hide'}
## The histogram graph of quantitative data in Dinesafe$Review shows that the data is normally distrubuted skewed to the left, where as Dinesafe$value shows that the data is not normally distrubuted.

## Histogram graph
hist(Dinesafe$REVIEW, col="GREEN")  
hist(Dinesafe$VALUE,col="YELLOW")   

```

```{r , results='hide'}
## Kernel Density Plots
Review <- density(Dinesafe$REVIEW)  
plot(Review)

```

```{r , results='hide'}
## Bar chart representation of a categorical values in Cusine Typek Inspection Result and Severity columns.

## CUISINE TYPE FREQUECY
Cuisin <- table(Dinesafe$CUISINE_TYPE)
barplot(Cuisin, main="Cuisine Type", ylab="Frequency", beside=TRUE, col = rainbow(17),las=2, horiz=FALSE)

## Food Hygiene Inspection Result
Inspection <- table(Dinesafe$ESTABLISHMENT_STATUS)
barplot(Inspection, main="Food Hygiene Inspection Result", xlab="Inspection Result", ylab="Establishment Count", col=c("red","yellow","green"), beside=TRUE)

## Food Hygiene Inspection Severity
Severity <- table(Dinesafe$SEVERITY)
barplot(Severity, main="Food Hygiene Inspection Severity", xlab="", ylab="Count", col=c("red","yellow","green","White","Orange"), beside=TRUE,las=2)

```

```{r , results='hide'}
## Boxplot of data distribution representation of Review and Value columns with its minimum, maximum, median, 1st quartile, 3rd quartile as well as outliers.

## The Review boxplot shows that the rating value range is quite close, ie between 2.5 and 4 with most of the data is concertrated above the median value of 3. Outlier data are above 4 and below 2.5 and mean value is 1.5
boxplot(Dinesafe$REVIEW,main = toupper("Boxplot of Review Column"),ylab = "Review Rating",col = "blue")
abline(h=mean(Dinesafe$VALUE, na.rm = T), lty=2)

## The Value boxplot shows that it has one outlier and the data distriubtion is between 1 and 2 and the median and mean values are close to each other around 1.5
boxplot(Dinesafe$VALUE,main = toupper("Boxplot of VALUE Column"),ylab = "Value",col = "green")
abline(h=mean(Dinesafe$VALUE, na.rm = T), lty=2)

```

5.2 Bivarient Data Analysis

5.2.1 - Mean & Standard deviation of Review and Value data against Estalishment status.

Based on the mean and standard deviation value of the review rating food premises that failed inspection had a mean review ratnig below the passed premises.  Also failed food premises has  a higher standard deviation value as compared to those who passed. 

On the other hand the relationship between mean/standard deviation value and inspection outcome is not observer due to consistent result across all three values.  
```{r  , results='hide'}
## Mean Review data against establishment inspection status
tapply(Dinesafe$REVIEW , Dinesafe$ESTABLISHMENT_STATUS, mean)
cat("\n")
## Standard Deviation of Review data against establishment inspection status
tapply(Dinesafe$REVIEW , Dinesafe$ESTABLISHMENT_STATUS, sd)  
cat("\n")
## Mean value data against establishment inspection status
tapply(Dinesafe$VALUE , Dinesafe$ESTABLISHMENT_STATUS, mean)
cat("\n")
## Standard Deviation of Value data against establishment inspection status
tapply(Dinesafe$VALUE , Dinesafe$ESTABLISHMENT_STATUS, sd)  
```

5.2.2 - Categorical data analysis against the numerical Review Rating column using bwplot
```{r , results='hide'}
library(lattice)
bwplot(ESTABLISHMENT_STATUS ~ REVIEW, data = Dinesafe, ylab = "Inspection Status", xlab = "Review Rating")
bwplot(SEVERITY ~ REVIEW, data = Dinesafe, ylab = "Inspection Outcome Severity", xlab = "Review Rating")
bwplot(CUISINE_TYPE ~ REVIEW, data = Dinesafe, ylab = "Cuisine Type", xlab = "Review Rating")

```
5.2.3 - Categorical data analysis against the numerical Value Rating column using bwplot
```{r , results='hide'}
## Categorical data analysis
bwplot(ESTABLISHMENT_STATUS ~ VALUE, data = Dinesafe, ylab = "Inspection Status", xlab = "Value Rating")
bwplot(SEVERITY ~ VALUE, data = Dinesafe, ylab = "Inspection Outcome Severity", xlab = "Value Rating ")
bwplot(CUISINE_TYPE ~ VALUE, data = Dinesafe, ylab = "Cuisine Type", xlab = "Value Rating")
```


5.2.4 - Categorical data analysis of District against the numerical data of Review and Value Rating column using qplot
```{r , results='hide'}
qplot(REVIEW, DISTRICT, data=Dinesafe) + xlim(0, 5)
qplot(VALUE, DISTRICT, data=Dinesafe) + xlim(0, 5)
```
5.2.5 - Relationship graphy between two categorical columns (INSPECTION STATUS AND SEVERITY)
```{r , results='hide'}
library(gmodels)
joint = CrossTable(Dinesafe$ESTABLISHMENT_STATUS, Dinesafe$SEVERITY)
joint$t
joint_count = joint$t
barplot(joint_count, beside = TRUE, col = c("Red","Yellow","Green"),ylab = "Frequency", xlab = "Severity")
legend("topright", c("Closed","Conditional Pass","Pass"), pch=15, col = c("Red","Yellow","Green"))

```

5.2.6 - Scattered graph showing a relatship between two numerical values
Based on the graph below there is no linear relationship between a restaurant review rating and value for money rating as the values are scattered all over the box and doesn't follow the simple linear regression model line.

```{r , results='hide'}
plot (Dinesafe$REVIEW, Dinesafe$VALUE,  col = c("RED", "BLUE"), xlab="REVIEW", ylab="VALUE", main="Review against Value graph", type = "b")
legend("topright", legend=c("Review", "Value"), col=c("green", "blue"), lty=1:2, cex=1, box.lty=2)
abline(lm(Dinesafe$REVIEW ~ Dinesafe$VALUE))


```




5.3 - Multivariant Data Analysis

5.3.1 - Plot of a simple scattered matrix between three columns (REVIEW, DISTRICT & ESTABLISHMENT TYPE) & (VALUE, DISTRICT & ESTABLISHMENT TYPE)
```{r , results='hide'}
pairs(~REVIEW+DISTRICT+ESTABLISHMENT_STATUS,data=Dinesafe,  main="Simple Scatterplot Matrix")
pairs( ~VALUE+DISTRICT+ESTABLISHMENT_STATUS,data=Dinesafe,  main="Simple Scatterplot Matrix")
```


5.3.2- Aggregation between multiple columns categorical and numerical values. 
```{r , results='hide'}
head(aggregate(Dinesafe$REVIEW ~ Dinesafe$ESTABLISHMENT_STATUS + Dinesafe$CUISINE_TYPE + Dinesafe$DISTRICT, FUN=mean),10)
head(aggregate(Dinesafe$REVIEW ~ Dinesafe$ESTABLISHMENT_STATUS + Dinesafe$CUISINE_TYPE, FUN=length),10)
```
# . Write final data frame to csv file. 
```{r}

write.csv(Dinesafe, file = "D:/CAPSTONE/CAPSTONE/DATASET/Final_Dinesafe.csv", row.names= TRUE)

```