---
title: "DineSafe"
author: "Mohammed Amir"
date: "February 18, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(Amelia)
```

## Load Dataset
```{r}
Dinesafe = read.csv("D:/CAPSTONE/data/DineSafe_02162017.csv")
Address  = read.csv("D:/CAPSTONE/data/FULL_ADDRESS_02262017.csv")

```

## Review dataset structure
```{r}
str(Dinesafe)
str(Address)
```

### Merge Dinesafe and Address dataset based on establishment id column
```{r}
Dinesafe <- merge(Dinesafe,Address,by="ESTABLISHMENT_ID")
```

## Remove Columns from data set 
```{r}
### Remove COURT_OUTCOME and AMOUNT_FINED Columns from Dinesafe dataset
Dinesafe <- subset(Dinesafe, select = -c(ROW_ID, COURT_OUTCOME,AMOUNT_FINED) )

### Remove Establishement Name from Address dataset to avoid duplicate rows 
Dinesafe <- subset(Dinesafe, select = -c(ESTABLISHMENT_NAME.y, ESTABLISHMENT_ADDRESS) )

## Rename ESTABLISHMENT_NAME.x column name to ESTABLISHMENT_NAME
colnames(Dinesafe)[colnames(Dinesafe) == 'ESTABLISHMENT_NAME.x'] <- 'ESTABLISHMENT_NAME'

```

## Dimension of Dinesafe
```{r}
dim(Dinesafe)
```

## Summary of Dinesafe
```{r}
summary(Dinesafe)
```


## Dataset Structure
```{r}
str(Dinesafe)
```



## Convert VALUE from factor to numeric Data Structure
```{r}
## use suppressWarnings to avoid getting "NAs introduced by coercion"
Dinesafe$VALUE <- suppressWarnings(as.numeric(as.character(Dinesafe$VALUE)))

```


## Convert INSPECTION_DATE from factor to Date Data Structure
```{r}
Dinesafe$INSPECTION_DATE <- dmy(as.character(Dinesafe$INSPECTION_DATE)) 
```


## Show Complete Case Rows with no missing (NA) value
```{r}
DinesafeCS  <- Dinesafe[complete.cases(Dinesafe),]
DinesafeCS
nrow(DinesafeCS)

```

## Show Rows with Missing (NA) value
```{r}
DinesafeNA  <- Dinesafe[!complete.cases(Dinesafe),]
DinesafeNA
nrow(DinesafeNA)
```

## Show missingness in Dinesafe dataset
```{r}

#Display and quantify missing values
apply(Dinesafe, 2, function(x) sum(is.na(x)))
 
# Plot missingness map
missmap(Dinesafe, col = c("Blue","Yellow"), y.cex = 0.8, x.cex = 0.8, legend = TRUE, rank.order = "False" ,main = "Dinesafe missingness map", y.labels = NULL,y.at = NULL)
```


## Impute Dinesafe$REVIEW with Mean Review Value for each Review based cuisine type
```{r}
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)

```



## Impute Dinesafe$VALUE with Mean Value for each based cuisine type
```{r}
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)

```

## Double Check Rows with Missing (NA) value after imputation
```{r}
DinesafeCS1  <- Dinesafe[complete.cases(Dinesafe),]
DinesafeCS1
nrow(DinesafeCS1)

DinesafeNA1  <- Dinesafe[!complete.cases(Dinesafe),]
DinesafeNA1
nrow(DinesafeNA1)

```
### Univariate Visualization     
```{r}
hist(Dinesafe$REVIEW, xlab = "Review Point", main="Histogram for Air Passengers", 
     border="blue", 
     col="YELLOW", 
     las=1, 
     breaks=5, 
     prob = TRUE)
lines(density(Dinesafe$REVIEW))
boxplot(Dinesafe$REVIEW)
plot(Dinesafe$REVIEW)
barplot(Dinesafe$REVIEW)
```

### Bivariate Visualization     
```{r}
plot (Dinesafe$INSPECTION_DATE,Dinesafe$REVIEW,  col = c("green", "red"), xlab="Inspection Date", ylab="VALUE", main="INSPECTION_DATE vs Review realtionship graphy")
legend("topright", legend=c("TotalFat", "Calories"), col=c("green", "red"), lty=1:2, cex=1, box.lty=1)


###barplot(Dinesafe$VALUE, Dinesafe$REVIEW, xlab = "Value", ylab = "Review")

```


### ### Multivariate Visualization  
```{r}

plot(Dinesafe)
````


```{r}
write.csv(total, file = "D:/CAPSTONE/data/Total.csv", row.names= TRUE)

````

