---
title: "DineSafe 1.1"
author: "Mohammed Amir"
date: "March 01, 2017"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate) ## Convert factor to date correctly
library(Amelia)    ## Missingness map
library(ggmap)
library(ggplot2)
```

## Load Dataset
```{r , results='hide'}
## Convert NULL values to NA
Dinesafe = read.csv("D:/CAPSTONE/data/DineSafe_02162017.csv", na.strings='NULL')
Address  = read.csv("D:/CAPSTONE/data/ADDRESS_02262017.csv", na.strings='NULL')

```

# Identify column names
```{r , results='hide'}
colnames(Dinesafe)

```

## Review dimension of dataset (Row by column)
```{r , results='hide'}
dim(Dinesafe)
dim(Address)
```


## Review summary of dinesafe dataset
```{r , results='hide'}
summary(Dinesafe)
```

## Review summary of Address dataset
```{r , results='hide'}
summary(Address)
```

## Review structure of Dinesafe dataset
```{r , results='hide'}
str(Dinesafe)
```

## Review structure of Address dataset
```{r , results='hide'}
str(Address)
```

### Merge Dinesafe and Address dataset based on establishment id column
```{r , results='hide'}
Dinesafe <- merge(Dinesafe,Address,by="ESTABLISHMENT_ID")
```

## Remove Columns from data set 
```{r , results='hide'}
### Remove COURT_OUTCOME and AMOUNT_FINED Columns from Dinesafe dataset
Dinesafe <- subset(Dinesafe, select = -c(ROW_ID, COURT_OUTCOME,AMOUNT_FINED,LONG_ADDRESS) )

### Remove Establishement Name from Address dataset to avoid duplicate rows 
Dinesafe <- subset(Dinesafe, select = -c(ESTABLISHMENT_NAME.y, ESTABLISHMENT_ADDRESS) )

## Rename ESTABLISHMENT_NAME.x column name to ESTABLISHMENT_NAME
colnames(Dinesafe)[colnames(Dinesafe) == 'ESTABLISHMENT_NAME.x'] <- 'ESTABLISHMENT_NAME'

```


## Convert VALUE from factor to numeric Data Structure
```{r , results='hide'}
## use suppressWarnings to avoid getting "NAs introduced by coercion" message
Dinesafe$VALUE <- suppressWarnings(as.numeric(as.character(Dinesafe$VALUE)))
```

## Convert INSPECTION_DATE from factor to Date Data Structure
```{r}
Dinesafe$INSPECTION_DATE <- dmy(as.character(Dinesafe$INSPECTION_DATE)) 
```

## Convert INSPECTION_DATE from factor to Date Data Structure
```{r , results='hide'}
Dinesafe$CUISINE_TYPE  <- as.character(Dinesafe$CUISINE_TYPE)
Dinesafe$ESTABLISHMENT_TYPE  <- as.character(Dinesafe$ESTABLISHMENT_TYPE)
Dinesafe$ESTABLISHMENT_STATUS <- as.character(Dinesafe$ESTABLISHMENT_STATUS) 
Dinesafe$SEVERITY <- as.character(Dinesafe$SEVERITY)
Dinesafe$ACTION <- as.character(Dinesafe$ACTION)
Dinesafe$ACTION <- as.character(Dinesafe$ACTION)
Dinesafe$ACTION <- as.character(Dinesafe$ACTION)

```


```{r , results='hide'}
str(Dinesafe)
```

## Complete Case Rows with no missing (NA) value
```{r , results='hide'}
Complete_Dinesafe  <- Dinesafe[complete.cases(Dinesafe),]
nrow(Complete_Dinesafe)

```

## Rows with Missing (NA) value
```{r , results='hide'}
Dinesafe_NA  <- Dinesafe[!complete.cases(Dinesafe),]
nrow(Dinesafe_NA)
```

## Show missingness in Dinesafe dataset
```{r , results='hide'}

#Quantify missing values
apply(Dinesafe, 2, function(x) sum(is.na(x)))
 
# Plot missingness map
missmap(Dinesafe, col = c("Red","Grey"), y.cex = 0.8, x.cex = 0.8, legend = TRUE, rank.order = "False" ,main = "Dinesafe missingness map", y.labels = NULL,y.at = NULL)
```

## Impute Dinesafe$REVIEW with Mean Review Value for each missing review value based cuisine type
```{r , results='hide'}
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$REVIEW[is.na(Dinesafe$REVIEW) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$REVIEW[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)

```



## Impute Dinesafe$VALUE with Mean Value for each missing value based cuisine type
```{r , results='hide'}
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="African"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="African"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bakeries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bakeries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Bar"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Bar"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Cafe"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Cafe"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Caribbean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Caribbean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Deli"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Deli"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Dessert"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Dessert"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="European"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="European"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Far Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Far Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Pastries"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Pastries"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="South East Asian"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="South East Asian"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Latin American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Latin American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Mediterranean"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Mediterranean"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Middle Eastern"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Middle Eastern"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="North American"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="North American"], na.rm=TRUE)
Dinesafe$VALUE[is.na(Dinesafe$VALUE) & Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"] = mean(Dinesafe$VALUE[Dinesafe$CUISINE_TYPE=="Juicery & Smoothies"], na.rm=TRUE)

```

```{r , results='hide'}
## Impute Severity column if it is NA and Establishment Status is PASS
Dinesafe$SEVERITY[is.na(Dinesafe$SEVERITY) & Dinesafe$ESTABLISHMENT_STATUS == "Pass"] = "N - No Action"

## Impute Action column if it is NA and Establishment Status is PASS & Severity is No Action
Dinesafe$ACTION[is.na(Dinesafe$ACTION) & Dinesafe$ESTABLISHMENT_STATUS == "Pass" & Dinesafe$SEVERITY == "N - No Action"] = "No Action Required"
```

## Double Check Rows with Missing (NA) value after imputation
```{r , results='hide'}
Complete_Dinesafe_1  <- Dinesafe[complete.cases(Dinesafe),]
Complete_Dinesafe_1
nrow(Complete_Dinesafe_1)

Dinesafe_NA1  <- Dinesafe[!complete.cases(Dinesafe),]
Dinesafe_NA1
nrow(Dinesafe_NA1)

```

### Univariate Analysis     
```{r , results='hide'}
hist(Dinesafe$REVIEW, xlab = "Review Rating", main="Histogram of Diner Review", 
  border="blue", col="GREEN", las=1, breaks=5,prob = TRUE)

hist(Dinesafe$VALUE, xlab = "Dollar Value", main="Histogram of Diner Value", 
     border="blue", col="YELLOW", las=1, breaks=5,prob = TRUE)
```

```{r  , results='hide'}
## CUISINE TYPE FREQUECY
status <- table(Dinesafe$CUISINE_TYPE)
barplot(status, main="Cuisine Type", ylab="Frequency", beside=TRUE, col = rainbow(17),las=2, horiz=FALSE)

## Food Hygiene Inspection Result
status <- table(Dinesafe$ESTABLISHMENT_STATUS)
barplot(status, main="Food Hygiene Inspection Result", xlab="Inspection Result", ylab="Establishment Count", col=c("red","yellow","green"), beside=TRUE)

## Food Hygiene Inspection Result
status1 <- table(Dinesafe$SEVERITY)
barplot(status1, main="Food Hygiene Inspection Severity", xlab="", ylab="Count", col=c("red","yellow","green","White","Orange"), beside=TRUE,las=2)
```

```{r , results='hide'}
boxplot(Dinesafe$REVIEW,
        main = toupper("Boxplot of Review"),
        ylab = "Review Rating",
        col = "blue")

boxplot(Dinesafe$VALUE,
        main = toupper("Boxplot of VALUE"),
        ylab = "Value",
        col = "green")


```

### Bivariate Analysis & Visualization 
```{r , results='hide'}
## Inspection outcome by Cuisine Stype
Inspection_Status_By_Cuisine_Type <- table(Dinesafe$CUISINE_TYPE, Dinesafe$ESTABLISHMENT_STATUS)
Inspection_Status_By_Cuisine_Type


##Scatter plot of restaurant review rating vs value
plot(Dinesafe$REVIEW, Dinesafe$VALUE, xlab="Review",ylab="Value",col=c("red","black"),cex=2) 

```